How is experience transformed into memory?

### 2018-08-30 paper/main.tex
substantial reorganization

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[font=small,labelfont=bf]{caption}
\usepackage{geometry}
\usepackage{natbib}
\usepackage{pxfonts}
\usepackage{graphicx}
\usepackage{newfloat}
\usepackage{setspace}
\DeclareFloatingEnvironment[name={Supplementary Figure},fileext=lof]{suppfigure}
\doublespacing

@@

\begin{figure}[t]
\centering
\includegraphics[width=1\textwidth]{figs/1_analysis_schematic_bw.pdf}
\caption{\small \textbf{Schematic of the analysis approach.} For each scene in the video, text descriptions were generated by hand. Three exemplary time points are displayed here.  Below the video descriptions are text samples from an example participant's verbal recall transcript.  We fit a topic model to the moment-by-moment video text descriptions and transformed participants' verbal recall transcripts using this same model. The bar charts display the resulting topic model weights for the video (in blue) and recall (in green) for three example topic dimensions.}
\label{fig:schematic}
\end{figure}

\subsection*{Modeling the naturalistic stimulus and verbal recall}
We fit a topic model \citep{BleiEtal03} to hand-annotated text descriptions of scenes from the video. The text descriptions contained details of the scene such as the characters present, location, and a short summary of the scene (see Fig.~\ref{fig:schematic} for example text). As depicted in Fig.~\ref{fig:model}A, the topic vectors are sparse and change slowly over time. Furthermore, there are clear transitions from one topic ``state'' to the next, possibly indexing scene transitions in the stimulus. To get a better handle on this temporal structure, we computed a timepoints (1976) by timepoints (1976) correlation matrix of the video model (Fig.~\ref{fig:model}C).  This correlation matrix reveals that the model has a strong, block-diagonal structure with very low off-diagonal correlation values, suggesting the representations of each block are unique and highly discriminable. We found that the ``narrative details'' (a few sentences summarizing the dialogue and specific happenings of the scene) of the video were the most important feature driving this model structure (see Supp. Fig.~\ref{fig:feature-importance}A).

After watching the video, participants verbally recalled (in order) as much of the episode as they could.  We used the same topic model (fit with the text descriptions of the video) to transform participants' verbal recall transcripts. An example participant's (\#13) recall model is plotted in Fig.~\ref{fig:model}B. Notably, participant \#13's recall model appears visually similar to the video model. Similar to the video model analysis (Supp. Fig.~\ref{fig:feature-importance}A), we found that the ``narrative details'' feature was most important for driving the video/recall relationship (Supp. Fig.~\ref{fig:feature-importance}B).  Like the video model, topic vectors were sparse and changed gradually.  Then, we investigated the temporal structure of the recall matrices. From each participant's recall model, we computed a sentences (68 to 294) by sentences (68 to 294) correlation matrix (Fig.~\ref{fig:model}D). Like the video model, each participant's recall correlation matrix exhibited a strong block-diagonal structure (Supp. Fig.~\ref{fig:corrmats}). Notably, this suggests that participants recounted the video in discriminable segments, likely related to the recall of specific events from the video.

\begin{figure}[t]
\centering
\includegraphics[width=\textwidth]{figs/2_eventseg.pdf}
\caption{\small \textbf{Modelling naturalistic stimuli and recall.} A depiction of our analysis pipeline. For all plots, darker colors indicate greater values and the range of each plot is 0-1.  \textbf{A.} A timepoints (1976) by topics (100) matrix representing the video stimulus.  Each row represents the most likely mixture of topics for a given timepoint (i.e., topic weights). Each column represents a different topic. \textbf{B.} A sentences (294) by topics (100) matrix representing participant \#13's recall. \textbf{C.} A viewing-time (1976) by viewing-time (1976) correlation matrix representing the correlation of each moment of the video model with every other moment of the video model. The yellow boxes represent `events' recovered by a hidden Markov model. \textbf{D.} A recall-time (294 sentences) by recall-time (294) correlation matrix for participant \#13. \textbf{E.} An events (34) by topics (100) matrix where each row represents the average topic vector for each event in the video model.  \textbf{F.} An events (27) by topics (100) matrix where each row represents the average topic vector for each event in participant \#13's recall model. \textbf{G.} A recall events (27) by video events (34) correlation matrix for participant \#13. The cells with a yellow border identify the video event with the highest correlation to a given recall event. \textbf{H.} A group averaged recall events (34) by video events (34) correlation matrix.  The cells with yellow borders are the video event with the highest correlation to a given average recall event.}
\label{fig:model}
\end{figure}

@@

Next, we created a video ``event model'' by averaging together neighboring topic vectors that the were classified to be in the same event, resulting in an events (34) by topics (100) matrix (Fig.~\ref{fig:model}E).  We performed the same procedure for the recall matrices (Fig.~\ref{fig:model}F). Then, we computed the correlation between video and recall event models, resulting in a video events (34) by recall events (8-27, depending on the participant) correlation matrix (Fig~\ref{fig:model}G for example, Supp. Fig~\ref{fig:matchmats} for all participants). These matrices represent the similarity (correlation) between each video event and each recall event (for each participant). To determine which video event a particular recall event was most likely describing, we found the index of the video event with the highest correlation to the recall event (i.e., the argmax).  This is depicted in Fig.~\ref{fig:model}G as the cells highlighted with a yellow border. Notably, our algorithm suggests that the example participant recalled most of the video events and did so in order.

@@

\begin{figure}[t]
\centering
\includegraphics[width=1\textwidth]{figs/3_trajectory.pdf}
\caption{\small \textbf{Video and recall trajectory plots.} \textbf{A.} 2-dimensional embedding of the video events model (reduced with UMAP algorithm). The arrows indicate the forward direction of the video events. \textbf{B.} 2-dimensional embedding of the average recall events model.  The colors refer to the most similar video events. The directional lines connecting the points represent the true video event order. The arrows represent the group-average direction of all recall event transitions (i.e., a line segment connecting two consecutive recall event vectors) that intersected a circle (radius = 0.5) centered on the origin of the arrow. \textbf{C.} Individual trajectory plots for each participant overlayed with the video model (black). The direction of the recall sequence is indicated by the direction of arrows connecting each recall event.  The colors refer to the most likely video event.}
\label{fig:trajectory}
\end{figure}

@@

\subsection*{Brain regions that mirror the correlational structure of naturalistic stimuli and memories}.
\begin{figure}[t]
\centering
\includegraphics[width=1\textwidth]{figs/4_brains_panels.pdf}
\caption{\small \textbf{Video and recall model correlational structure searchlight analysis.} \textbf{A.} The correlation matrix representing the temporal structure of the video model. Below it is a surface plot of brain regions whose activity patterns mirrored the correlational structure of video model during viewing.  \textbf{B.} The left matrix is S13's video-recall correlation matrix before warping. The yellow stars represent the path recovered by dynamic time warping. The right matrix is the result after warping. The same warping path was applied to the fMRI timeseries (separately for each participant). A surface plot of brain regions whose activity patterns mirrored the participant-specific recall correlation matrix. Significance threshold (p$<$.05) for both video and recall analyses derived from a permutation procedure by phase shifting the video model and recomputing the z-scores.}
\label{fig:brainz}
\end{figure}

@@

While each participant remembered roughly the same sequence of events, there were subtle idiosyncracies in the way each individual described the events. For example, the event-by-event similarity structure (i.e., the off-diagonal values of the recall correlation matrices) differed across individuals (Supp. Fig.~\ref{fig:corrmats}). We hypothesized that idiosyncracies in the recall dynamics across participants might be related to the temporal structure of brain networks representing the events during video viewing. In this analysis, we identified brain areas whose responses at the time of viewing exhibited the same (segmented) idiosyncratic dynamics that later appeared in that participants’ recalls. Using a procedure similar to the approach above (see Methods for details), we searched for regions whose correlational structure during viewing was related to the idiosyncratic structure of the participant's recall model. The resulting map suggests that the correlational structure of a network of brain regions including the ventromedial prefrontal cortex, anterior cingulate, and right medial temporal lobe mirror the idiosyncratic structure of participants' recalls (Fig.~\ref{fig:brainz}).

\begin{figure}[t]
\centering
\includegraphics[width=1\textwidth]{figs/5_topics.pdf}
\caption{\small \textbf{Analysis of topics by event memorability.} \textbf{A.} Group-averaged correlation between each video event and closest recall event. Error bars represent 95\% confidence intervals. Stars indicate top (green) and bottom (red) 3 remembered events across participants \textbf{B.} Wordles (top 200 words) representing a weighted average of video event topic vectors weighted by event memorability (e.g., the correlation values in A). The top wordle (green) contains words from the most remembered events and the bottom (red) contains words from least memorable events. The word sizes are proportional to the word's ``activation''. \textbf{C.} Trajectory represents the video event model embedded in a 2D space using UMAP. The large colored dots are video events. The small colored dots are individual recall events across all participants, where the colors correspond to the closest video event. The circular wordles represent the top 3 most (green circles) and least (red) memorable video events.  The left side of the wordle circles contain words associated with the video event vector and the right side contains words associated with the average recall event vector.}
\label{fig:topics}
\end{figure}

@@

\subsubsection*{Extracting events using a hidden Markov model}
The topic model timepoint-by-timepoint correlation matrices all exhibited a block-diagonal structure (with small off-diagonal values), suggesting that the models were comprised of a number of sequential `states' (or events, see Supp. Fig.~\ref{fig:corrmats}). To capture this structure, we fit the video and each recall model using a hidden Markov model (HMM). Given a number of states or events ($k$), the HMM recovers a set of labels that segments consecutive timepoints into $k$ events ~\citep{Rabi89, BaldEtal17}. To implement this analysis, we used the Brainiak toolbox \citep{BaldEtal17, Brainiak}.

@@

\subsubsection*{Matching recall events to video events}
To estimate which video event each recall event referred to, we correlated the video events model and each recall events model. This resulted in a video events (34) by recall events (8-27) correlation matrix (for each participant) which contains the similarity between each video event and each recall event (see Supp. Fig.~\ref{matchmats}).  To find the most likely video event that a given recall event referred to, we computed the argmax over the columns of this matrix.  The result was a list of video event indices for each participant. These indices are analogous to the values found in a ``recall matrix'' from a free recall list learning experiment, but represent the recall of particular events (instead of words, for example).

@@

% SUPPLEMENTAL MATERIALS

\pagebreak
\begin{center}
\textbf{\large Supplemental Materials: How is experience transformed into memory?}
\end{center}
\setcounter{equation}{0}
\setcounter{figure}{0}
\setcounter{table}{0}
\setcounter{page}{1}
\setcounter{section}{0}
\makeatletter
\renewcommand{\theequation}{S\arabic{equation}}
\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\bibnumfmt}[1]{[S#1]}
\renewcommand{\citenumfont}[1]{S#1}

% Supplemental materials
\begin{suppfigure}[t]
\centering
\includegraphics[width=1\textwidth]{figs/supp1_corrmats.pdf}
\caption{\small \textbf{Recall model correlation matrices and event segmentation fits.} Each participant's timepoint-by-timepoint recall correlation matrix.  The yellow boxes represent ``events'' identified by a hidden Markov model.}
\label{fig:corrmats}
\end{suppfigure}

\begin{suppfigure}[t]
\centering
\includegraphics[width=1\textwidth]{figs/supp2_matchmats.pdf}
\caption{\small \textbf{Video-recall event model correlation matrices.} Each participant's video event by recall event correlation matrix.  The yellow boxes represent the maximum correlation in each column.}
\label{fig:matchmats}
\end{suppfigure}

\begin{suppfigure}[t]
\centering
\includegraphics[width=1\textwidth]{figs/supp4_feature_value.pdf}
\caption{\small \textbf{Impact of individual features on topic modeling analysis.} \textbf{A.} Contribution of each feature to model structure. Bars represent the correlation of a video model trained in the absence of a given feature to the model trained on all features. \textbf{B.} Contribution of each feature to video model/recall model relationship. The leftmost bar represents the across-participants mean correlation between the video model and recall models trained on all features. Subsequent bars represent the same relationship between video and recall models trained in the absence of a given feature. Error bars are the standard error of the mean across participants. \textbf{C.}  Individual feature trajectory similarity matrix. The first row/column is the topic trajectory the full video model. Each subsequent row/column is that of a single feature. Shading corresponds to the value of the correlation coefficient (Pearson's $r$) .}
\label{fig:feature-importance}
\end{suppfigure}

\section*{Naturalistic extensions of classic list-learning analyses}
Just like in a traditional free recall list-learning experiment where participants view a list of words and then verbally recall them, our video-recall matching analysis approach affords us the ability to analyze memory in the same way. The recalled events can be treated as ``items'' analogous to words recalled in a list-learning study. Here, we sought to characterize memory performance/dynamics by extending classic analyses originally designed for list-learning experiments to more naturalistic settings.

First, we asked whether the estimated number of recall events ($k$) by participant was related to hand-annotated accuracy as published in Chen et al. (2017).  We found a strong positive correlation where participants with a greater number of recall events also had better overall memory performance (Pearson's $r(16) = 0.67, p = 0.003$). Then, we considered how participants initiated the recall sequence (known in the literature as the `probability of first recall' or `PFR'). We found that participants tended to initiate their recall sequences with the first few events (Supp. Fig.~\ref{fig:list-learning}A), which is qualitatively very similar to previously published list learning experiments ~\citep{HowaKaha99}. Next, we considered another well-studied memory measure in the list-learning literature, the lag conditional response probability curve (or lag-CRP) ~\citep{Kaha96}. The result suggests a strong bias to transition sequentially events in the forward direction (Supp. Fig.~\ref{fig:list-learning}B). Finally, we assessed memory performance for each event in the video as a function of its serial position during encoding (Supp. Fig.~\ref{fig:list-learning}C). We did not observe the classic ``primacy'' and ``recency'' pattern which is prevalent in the literature ~\citep{Murd62a}. We also considered two additional across-participant measures of recall that characterize memory organization: temporal clustering and semantic clustering. We found that participants who clustered in time also recalled a greater number of events (Pearson's $r(16) = 0.62, p = 0.007$). Next, we assessed semantic clustering. We found that the semantic clustering score was related to memory performance across participants (Pearson's $r(16) = 0.55, p = 0.02$).  Thus, participants who organized their recalls with respect to the semantic information contained in the scene had better memory performance.

\section*{Additional measures of naturalistic memory}
To quantify the similarity between the video model and individual recall models, we considered a number of novel metrics.  First, we tested whether each participant's recall model matched the video model in a general sense. To do this, for each participant we filtered the video model to only include the events that the participant remembered and computed the root mean squared difference (RMSD) between the video model and the recall model. As an example, if the participant remembered all the events in order (with perfect precision), the expected distance value would be 0. However, if they remembered a subset of events, events out of order, or with low precision, the expected distance would be greater than 0. To assess significance, we performed a permutation test where we circularly shifted the recall model (10000 times) and recomputed the RMSD. The recall model significantly matched the video model for nine of the participants ($p < 0.05$, participants: 3-4, 8-13, 17 and the p-value for the rest of the participants was less than .25). Furthermore, the RMSD values were significantly correlated to hand annotated memory performance across participants (Pearson's $r(16) =-.57, p = 0.016$). Thus, a closer match between the video and recall event models was related to better recall performance.

Next, we tested whether participants who recalled more events were also more precise in their recollections. For each participant, we computed the correlation between each recall event and its matching video event (only for the events which they recalled). This resulted in a single number for each recalled event indexing how similar the recall event was to its matching video event (i.e the ``precision'' of the recall). We then averaged the correlations within participant. In line with our prediction, there was a strong correlation between hand annotated memory performance and precision suggesting that participants who remembered more events also remembered them more veridically (Pearson's $r(16) = 0.74, p = 0.0006$).

Then, we considered the distinctiveness of each recall event. That is, how uniquely a recall event matched a given video event compared to all other video events. We hypothesized that participants with high memory performance might describe each event in a more distinctive way (relative to those with lower memory performance who might describe events in a more general way). To this end, we computed a `distinctiveness' score for each participant (i.e., 1 - the correlation between a recall event and all non-matching video events).  Then, we averaged this measure over recall events within participant.  We found that participants with higher hand annotated memory performance also had higher distinctiveness scores (Pearson's $r(16) = 0.8, p = 0.0001$).

Lastly, we tested whether participants with better memory performance were also more likely to remember the events in order.  For each participant, we computed the Spearman rank correlation between the order of events that the participant recalled and the actual order of events (filtering events that were actually recalled).  We found that participants who recalled more events also recalled more of them in order (Pearson's $r(16) = 0.5, p = 0.04$). In summary, we found that better memory performance was associated with more precise, distinctive and ordered recalls.

\begin{suppfigure}[t]
\centering
\includegraphics[width=1\textwidth]{figs/supp3_list_learning.pdf}
\caption{\small \textbf{Naturalistic extensions of classic list-learning memory analyses.} A). The probability of first recall as a function of the serial position of the event during encoding. B). A lag-conditional response probability curve. Given recall of event i, the probability that the next recalled item will be from serial position i +/- lag. C). Proportion of events recalled as a function of serial position. All error bars are the standard error of the mean derived from a bootstrap resampling procedure.}
\label{fig:list-learning}
\end{suppfigure}

\section*{Supplemental Methods}
\subsection*{Quantifying the importance of features}
To determine the contribution of each feature to the structure of the video model, we examined the similarity between the temporal structure of models trained in the absence of a single feature and that of our original (i.e., full) model (Supp. Fig.~\ref{fig:feature-importance}A). First, we iteratively removed one transcribed feature from the scene descriptions and constructed a timepoints (1976) by topis (100) matrix using a topic model fit to the remaining features. We then represented the original model as well as the new model’s temporal structure as a timepoints-by-timepoints correlation matrix. Finally, we vectorized these correlation matrices and correlated them to each other resulting in a single number (for each feature removed from the model) representing the similarity of the ``feature-removed'' models to the full model.

In order to ascertain which features were important in relating the recall models to the video model, we similarly compared to the temporal structures of video and recall models deprived of a single feature at a time (Supp. Fig.~\ref{fig:feature-importance}B). For each feature removed, we transformed each participant’s recall transcript using a model trained on the feature-deprived video text windows (of 50 scene segments), and resampled the recall timeseries to match the shape of the video model (1976 timepoints). We then represented the temporal structures of each participant’s recall model as timepoints-by-timepoints correlation matrices and computed the average correlation with the temporal structure of the video model (across participants).

The two prior analyses examine information (about either the video model structure or the video/recall model relationship) lost in the absence of each feature. However, they do not consider redundancy in the information each feature contributes (e.g., how much unique information does "Character speaking" provide that "Character in focus" does not?). As a measure of information overlap between features, we computed the similarity between the topic trajectories of each individual annotated feature in addition to the full feature set. First, we singularly transformed the text of each feature using a model trained on the collection of all features. We then represented the single feature's model's temporal structure as a timepoints-by-timepoints correlation matrix, and compared the temporal structures pairwise by constructing a features-by-features correlation matrix ((Supp. Fig.~\ref{fig:feature-importance}C). We additionally computed the correlation between the individual features' temporal structures and that of the full video model as a measure of the proportion of information discernible from that feature alone.

\subsection*{List-learning analyses}
\textbf{Overall Accuracy.}  To get an overall measure of the quantity of information recalled, we computed the proportion of sucessfully recalled events by counting the number of unique recall events identified by the HMM model and dividing by the total number of video events.  We performed this analysis for each participant separately.

\textbf{Probability of first recall (PFR).}  The (PFR) analysis represents the probability that an item will be recalled first as a function of its serial position during encoding. We initialized a \# of participants (17) by \# of video events (34) matrix. Then for each participant, we found the index of the video event that was recalled first and filled in that index in the matrix with a 1.  Finally, we averaged over the rows of the matrix, resulting in a 1 by 34 array representing the proportion of participant that recalled an event as a function of serial position during encoding.

\textbf{Lag conditional probability curve (lag-CRP).} The lag-CRP represents the probability that the next item recalled will be of lag $i$ from the just recalled item. For each recall transition, we computed the lag between the current recall event and the next recall event, normalizing by the total number of possible transitions.  This resulted in a \# of participants (17) by lags (-33:+33) matrix. We averaged over the rows of this matrix to get a group-averaged lag-CRP.

\textbf{Serial position curve (SPC).} The SPC represents the proportion of participants that remember an item as a function of its serial position during encoding. We initialized a \# of participants (17) by \# of video events (34) matrix. Then, for each recall event (and each participant), we found the index of the video event that was recalled and filled it in with a 1. This resulted in a matrix where 1s indicate the successful recall of an event in serial position $n$ and zeros indicate the lack of recall for that event.  Lastly, we averaged over the rows of the matrix to get a 1 by 34 array representing the proportion of participants that recalled an event as a function of its serial position.

\textbf{Temporal clustering.} Temporal clustering measures the extent to which participants group their recall responses according to encoding position ~\citep{PolyEtal09}. For instance, if the participant recalled each item in the presentation order, this would result in a score of 1. If the participant recalled randomly with respect to presentation order, this would result in a score of ~.5.  For each event transition (and separately for each participant), we computed the rank similarity (euclidean distance) between the presentation position  of the current and next recall events. The scores were then averaged within participant to get a single number representing the extent of temporal clustering exhibited by a given participant.

\textbf{Semantic clustering.} Similar to temporal clustering, semantic clustering measures the extent to which participants group their recall responses according to semantic similarity ~\citep{PolyEtal09}. Here, we are using the topic vectors for each event as a proxy for its semantic content. Thus, similarity between the semantic content for two events can be computed by correlating their respective topic vectors.  For instance, if each consecutive recall was the next most similar event (in terms of its s), this would result in a score of ~1. If the participant recalled randomly with respect to semantic similarity, this would result in a score of ~.5.  For each event transition (and separately for each participant), we computed the rank similarity (correlation distance) between the current recall event and the next recall event. The scores were then averaged within participant to get a single number representing the extent of semantic clustering exhibited by a given participant.

\subsection*{Additional measures of naturalistic memory}
\textbf{Precision.} This measure gives us an indication of the specific match between a video event and recall event, where values approaching 1 are highly precise and lower values are inprecise.  We defined ``precision'' as the correlation between a recall event and its matching (i.e., argmax) video event.

\textbf{Distinctiveness.} Distinctiveness quantifies how similar a recall event is to all non-matching video events.  It provides a metric of how uniquely a particular recall event describes a particular video event.  To compute it, a given recall event is correlated to all video events, the argmax is removed and the rest of the values are averaged. The resulting value is subtracted from 1 such that larger values indicate a more distinctive recall event.

>>>>

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[font=small,labelfont=bf]{caption}
\usepackage{geometry}
\usepackage{natbib}
\usepackage{pxfonts}
\usepackage{graphicx}
\usepackage{newfloat}
\usepackage{setspace}
%\doublespacing

@@

\begin{figure}[tp]
\centering
\includegraphics[width=1\textwidth]{figs/1_analysis_schematic_bw.pdf}
\caption{\small \textbf{Schematic of the analysis approach.} For each scene in the video, text descriptions were generated by hand. Three exemplary time points are displayed here.  Below the video descriptions are text samples from an example participant's verbal recall transcript.  We fit a topic model to the moment-by-moment video text descriptions and transformed participants' verbal recall transcripts using this same model. The bar charts display the resulting topic model weights for the video (in blue) and recall (in green) for three example topic dimensions.}
\label{fig:schematic}
\end{figure}

\subsection*{Modeling the naturalistic stimulus and verbal recall}
We fit a topic model \citep{BleiEtal03} to hand-annotated text descriptions of scenes from the video. The text descriptions contained details of the scene such as the characters present, location, and a short summary of the scene (see Fig.~\ref{fig:schematic} for example text). As depicted in Fig.~\ref{fig:model}A, the topic vectors are sparse and change slowly over time. Furthermore, there are clear transitions from one topic ``state'' to the next, possibly indexing scene transitions in the stimulus. To get a better handle on this temporal structure, we computed a timepoints (1976) by timepoints (1976) correlation matrix of the video model (Fig.~\ref{fig:model}C).  This correlation matrix reveals that the model has a strong, block-diagonal structure with very low off-diagonal correlation values, suggesting the representations of each block are unique and highly discriminable. We found that the ``narrative details'' (a few sentences summarizing the dialogue and specific happenings of the scene) of the video were the most important feature driving this model structure (see Fig.~S3A).

After watching the video, participants verbally recalled (in order) as much of the episode as they could.  We used the same topic model (fit with the text descriptions of the video) to transform participants' verbal recall transcripts. An example participant's (\#13) recall model is plotted in Fig.~\ref{fig:model}B. Notably, participant \#13's recall model appears visually similar to the video model. Similar to the video model analysis (Fig.~S3A), we found that the ``narrative details'' feature was most important for driving the video/recall relationship (Fig.~S3B).  Like the video model, topic vectors were sparse and changed gradually.  Then, we investigated the temporal structure of the recall matrices. From each participant's recall model, we computed a sentences (68 to 294) by sentences (68 to 294) correlation matrix (Fig.~\ref{fig:model}D). Like the video model, each participant's recall correlation matrix exhibited a strong block-diagonal structure (Fig.~S1). Notably, this suggests that participants recounted the video in discriminable segments, likely related to the recall of specific events from the video.

\begin{figure}[tp]
\centering
\includegraphics[width=\textwidth]{figs/2_eventseg.pdf}
\caption{\small \textbf{Modelling naturalistic stimuli and recall.} A depiction of our analysis pipeline. For all plots, darker colors indicate greater values and the range of each plot is 0-1.  \textbf{A.} A timepoints (1976) by topics (100) matrix representing the video stimulus.  Each row represents the most likely mixture of topics for a given timepoint (i.e., topic weights). Each column represents a different topic. \textbf{B.} A sentences (294) by topics (100) matrix representing participant \#13's recall. \textbf{C.} A viewing-time (1976) by viewing-time (1976) correlation matrix representing the correlation of each moment of the video model with every other moment of the video model. The yellow boxes represent `events' recovered by a hidden Markov model. \textbf{D.} A recall-time (294 sentences) by recall-time (294) correlation matrix for participant \#13. \textbf{E.} An events (34) by topics (100) matrix where each row represents the average topic vector for each event in the video model.  \textbf{F.} An events (27) by topics (100) matrix where each row represents the average topic vector for each event in participant \#13's recall model. \textbf{G.} A recall events (27) by video events (34) correlation matrix for participant \#13. The cells with a yellow border identify the video event with the highest correlation to a given recall event. \textbf{H.} A group averaged recall events (34) by video events (34) correlation matrix.  The cells with yellow borders are the video event with the highest correlation to a given average recall event.}
\label{fig:model}
\end{figure}

@@

Next, we created a video ``event model'' by averaging together neighboring topic vectors that the were classified to be in the same event, resulting in an events (34) by topics (100) matrix (Fig.~\ref{fig:model}E).  We performed the same procedure for the recall matrices (Fig.~\ref{fig:model}F). Then, we computed the correlation between video and recall event models, resulting in a video events (34) by recall events (8-27, depending on the participant) correlation matrix (Fig~\ref{fig:model}G for example, Fig.~S2 for all participants). These matrices represent the similarity (correlation) between each video event and each recall event (for each participant). To determine which video event a particular recall event was most likely describing, we found the index of the video event with the highest correlation to the recall event (i.e., the argmax).  This is depicted in Fig.~\ref{fig:model}G as the cells highlighted with a yellow border. Notably, our algorithm suggests that the example participant recalled most of the video events and did so in order.

@@

\begin{figure}[tp]
\centering
\includegraphics[width=1\textwidth]{figs/3_trajectory.pdf}
\caption{\small \textbf{Video and recall trajectory plots.} \textbf{A.} 2-dimensional embedding of the video events model (reduced with UMAP algorithm). The arrows indicate the forward direction of the video events. \textbf{B.} 2-dimensional embedding of the average recall events model.  The colors refer to the most similar video events. The directional lines connecting the points represent the true video event order. The arrows represent the group-average direction of all recall event transitions (i.e., a line segment connecting two consecutive recall event vectors) that intersected a circle (radius = 0.5) centered on the origin of the arrow. \textbf{C.} Individual trajectory plots for each participant overlayed with the video model (black). The direction of the recall sequence is indicated by the direction of arrows connecting each recall event.  The colors refer to the most likely video event.}
\label{fig:trajectory}
\end{figure}

@@

\subsection*{Brain regions that mirror the correlational structure of naturalistic stimuli and memories}.
\begin{figure}[tp]
\centering
\includegraphics[width=1\textwidth]{figs/4_brains_panels.pdf}
\caption{\small \textbf{Video and recall model correlational structure searchlight analysis.} \textbf{A.} The correlation matrix representing the temporal structure of the video model. Below it is a surface plot of brain regions whose activity patterns mirrored the correlational structure of video model during viewing.  \textbf{B.} The left matrix is S13's video-recall correlation matrix before warping. The yellow stars represent the path recovered by dynamic time warping. The right matrix is the result after warping. The same warping path was applied to the fMRI timeseries (separately for each participant). A surface plot of brain regions whose activity patterns mirrored the participant-specific recall correlation matrix. Significance threshold (p$<$.05) for both video and recall analyses derived from a permutation procedure by phase shifting the video model and recomputing the z-scores.}
\label{fig:brainz}
\end{figure}

@@

While each participant remembered roughly the same sequence of events, there were subtle idiosyncracies in the way each individual described the events. For example, the event-by-event similarity structure (i.e., the off-diagonal values of the recall correlation matrices) differed across individuals (Fig.~S1). We hypothesized that idiosyncracies in the recall dynamics across participants might be related to the temporal structure of brain networks representing the events during video viewing. In this analysis, we identified brain areas whose responses at the time of viewing exhibited the same (segmented) idiosyncratic dynamics that later appeared in that participants’ recalls. Using a procedure similar to the approach above (see Methods for details), we searched for regions whose correlational structure during viewing was related to the idiosyncratic structure of the participant's recall model. The resulting map suggests that the correlational structure of a network of brain regions including the ventromedial prefrontal cortex, anterior cingulate, and right medial temporal lobe mirror the idiosyncratic structure of participants' recalls (Fig.~\ref{fig:brainz}).

\begin{figure}[tp]
\centering
\includegraphics[width=1\textwidth]{figs/5_topics.pdf}
\caption{\small \textbf{Analysis of topics by event memorability.} \textbf{A.} Group-averaged correlation between each video event and closest recall event. Error bars represent 95\% confidence intervals. Stars indicate top (green) and bottom (red) 3 remembered events across participants \textbf{B.} Wordles (top 200 words) representing a weighted average of video event topic vectors weighted by event memorability (e.g., the correlation values in A). The top wordle (green) contains words from the most remembered events and the bottom (red) contains words from least memorable events. The word sizes are proportional to the word's ``activation''. \textbf{C.} Trajectory represents the video event model embedded in a 2D space using UMAP. The large colored dots are video events. The small colored dots are individual recall events across all participants, where the colors correspond to the closest video event. The circular wordles represent the top 3 most (green circles) and least (red) memorable video events.  The left side of the wordle circles contain words associated with the video event vector and the right side contains words associated with the average recall event vector.}
\label{fig:topics}
\end{figure}

@@

\subsubsection*{Extracting events using a hidden Markov model}
The topic model timepoint-by-timepoint correlation matrices all exhibited a block-diagonal structure (with small off-diagonal values), suggesting that the models were comprised of a number of sequential `states' (or events, see Fig.~S1). To capture this structure, we fit the video and each recall model using a hidden Markov model (HMM). Given a number of states or events ($k$), the HMM recovers a set of labels that segments consecutive timepoints into $k$ events ~\citep{Rabi89, BaldEtal17}. To implement this analysis, we used the Brainiak toolbox \citep{BaldEtal17, Brainiak}.

@@

\subsubsection*{Matching recall events to video events}
To estimate which video event each recall event referred to, we correlated the video events model and each recall events model. This resulted in a video events (34) by recall events (8-27) correlation matrix (for each participant) which contains the similarity between each video event and each recall event (see Fig.~S2).  To find the most likely video event that a given recall event referred to, we computed the argmax over the columns of this matrix.  The result was a list of video event indices for each participant. These indices are analogous to the values found in a ``recall matrix'' from a free recall list learning experiment, but represent the recall of particular events (instead of words, for example).

@@
